Qualitative Probabilistic Programming

### 2015-06-05 paper/paper.tex
syntax highlight

\documentclass{article}
\usepackage{cite}
\usepackage{amsmath}
% \usepackage{minted}
\usepackage{hyperref}
\usepackage{amsfonts}
\usepackage{multicol} \usepackage{fancyheadings} \usepackage{pdfpages}
\usepackage{nips15submit_e,times}
\setlength{\emergencystretch}{10em}

@@

  Here is an example of a Quipp program to cluster 2d points into 2 clusters:
  \begin{verbatim}
var n = 100;

var Cluster = Categorical(2);
var Point = Vector(2, Double);
var getPoint = randFunction(Cluster, Point);
return function() {
  repeat(n, function(i) {
    var cluster = randomValue(Cluster);
    observe(getPoint, cluster);
  });
};
  \end{verbatim}

@@

  The first cluster is at (6.3, 5.0) and the second is at (5.0, 1.6).  They both
  have a standard deviation of 0.54 in the x direction and 0.69 in the y direction.
  We could use these parameters to fill in the generative model:
  \begin{verbatim}
    return repeat(n, function(i) {
      var cluster = randomInteger(nclusters);
      return [gaussian(cluster == 0 ? 6.3 : 5.0, 0.54),
              gaussian(cluster == 0 ? 5.0 : 1.6, 0.69)];
    });
  \end{verbatim}
  This model is estimated to assign probability density $e^{-393}$ to the data, yielding a perplexity of $e^{-393/150} = 0.0728$.

  Note that, because the two clusters are forced to have the same variance, they do not fit the data well,
  since the data has a different shape in each location.  To fix this problem, we can substitute the following
  model, which uses a separate random function for each cluster:
  \begin{verbatim}
var Cluster = Categorical(2);
var Point = Vector(2, Double);
var getPointFunctions = [randFunction(Point), randFunction(Point)];
return repeat(n, function(i) {
  var cluster = randomValue(Cluster);
  observe(getPointFunctions[cluster]);
});
\end{verbatim}
  Using this model, we get the following clusters:
  \begin{center}
    \includegraphics[scale=0.5]{../plots/irisclusters_indep.png}
  \end{center}
  This model (with the parameters filled in) is estimated to assign probability density $e^{-337}$ to the data, yielding a perplexity of 0.1058.  This is a significantly improved fit.

@@

\begin{verbatim}

@@

return function() {
  return repeat(n, function(i) {
    var cluster = randomInteger(nclusters);
    var x = observe(getX[cluster]);
    var y = observe(getY[cluster], x);
  });
};
\end{verbatim}

@@

  \subsection{Clustering}
{\small
\begin{verbatim}
var nclusters = 3;
var dim = 2;

@@

var Point = Vector(3, Double);
var Cluster = Categorical(3);
var getPoint = randFunction(Cluster, Point);
return function() {
  return repeat(n, function(i) {
    var cluster = randomValue(Cluster);
    observe(getPoint, cluster);
  });
};
\end{verbatim}
}

@@

{\small
\begin{verbatim}
var n = 50;

@@

return function() {
  repeat(n, function(i) {
    var whichClass = randomValue(Class);
    observe(classFeatures, whichClass);
  });
};
\end{verbatim}
}

@@

  \subsection{Factor analysis}
{\small
\begin{verbatim}

@@

return function() {
  return repeat(n, function(i) {
    var factors = randomValue(Factors);
    observe(getPoint, factors);
  });
};
\end{verbatim}
}

@@

  \subsection{Hidden Markov model}
{\small
\begin{verbatim}

@@

return function() {
  return repeat(n, function(sampIndex) {
    observeStates(randomValue(State), 0);
  });
};
\end{verbatim}
}

@@

\subsection{Latent Dirichlet allocation}
{\small
\begin{verbatim}

@@

\end{verbatim}
}

@@

{\small
\begin{verbatim}

@@

return function() {
  return repeat(inputs.length, function(sampIndex) {
    var hiddenLayer = inputToHidden(inputs[sampIndex]);
    observe(hiddenToOutput, hiddenLayer);
  });
};
\end{verbatim}
}

>>>>

\documentclass{article}
\usepackage{cite}
\usepackage{listings}
\usepackage{amsmath}
% \usepackage{minted1}
\usepackage{hyperref}
\usepackage{amsfonts}
\usepackage{multicol} \usepackage{fancyheadings} \usepackage{pdfpages}
\usepackage{nips15submit_e,times}
\setlength{\emergencystretch}{10em}

@@

  Here is an example of a Quipp program to cluster 2d points into 2 clusters:
\lstdefinelanguage{JavaScript}{
      morekeywords={break, case, catch, continue, debugger, default, delete,         do, else, false, finally, for, function, if, in, instanceof, new, null, return, switch, this, throw, true, try, typeof, var, void, while, with},
      morecomment=[s]{/*}{*/},
      morecomment=[l]//,
      morestring=[b]",
      morestring=[b]'
    }
  \lstset{language=JavaScript, xleftmargin=0.5in}

  \begin{lstlisting}
var n = 100;

var Cluster = Categorical(2);
var Point = Vector(2, Double);
var getPoint = randFunction(Cluster, Point);
return function() {
  repeat(n, function(i) {
    var cluster = randomValue(Cluster);
    observe(getPoint, cluster);
  });
};
  \end{lstlisting}

@@

  The first cluster is at (6.3, 5.0) and the second is at (5.0, 1.6).  They both
  have a standard deviation of 0.54 in the x direction and 0.69 in the y direction.
  We could use these parameters to fill in the generative model:
  \begin{lstlisting}
return repeat(n, function(i) {
  var cluster = randomInteger(nclusters);
  return [gaussian(cluster == 0 ? 6.3 : 5.0, 0.54),
          gaussian(cluster == 0 ? 5.0 : 1.6, 0.69)];
});
  \end{lstlisting}
  This model is estimated to assign probability density $e^{-393}$ to the data, yielding a perplexity of $e^{-393/150} = 0.0728$.

  Note that, because the two clusters are forced to have the same variance, they do not fit the data well,
  since the data has a different shape in each location.  To fix this problem, we can substitute the following
  model, which uses a separate random function for each cluster:
  \begin{lstlisting}
var Cluster = Categorical(2);
var Point = Vector(2, Double);
var getPointFunctions = [randFunction(Point), randFunction(Point)];
return repeat(n, function(i) {
  var cluster = randomValue(Cluster);
  observe(getPointFunctions[cluster]);
});
\end{lstlisting}
  Using this model, we get the following clusters:
  \begin{center}
    \includegraphics[scale=0.5]{../plots/irisclusters_indep.png}
  \end{center}
  This model (with the parameters filled in) is estimated to assign probability density $e^{-337}$ to the data, yielding a perplexity of 0.1058.  This is a significantly improved fit.

@@

\begin{lstlisting}

@@

return function() {
  return repeat(n, function(i) {
    var cluster = randomInteger(nclusters);
    var x = observe(getX[cluster]);
    var y = observe(getY[cluster], x);
  });
};
\end{lstlisting}

@@

  \subsection{Clustering}
{\small
\begin{lstlisting}
var nclusters = 3;
var dim = 2;

@@

var Point = Vector(3, Double);
var Cluster = Categorical(3);
var getPoint = randFunction(Cluster, Point);
return function() {
  return repeat(n, function(i) {
    var cluster = randomValue(Cluster);
    observe(getPoint, cluster);
  });
};
\end{lstlisting}
}

@@

{\small
\begin{lstlisting}
var n = 50;

@@

return function() {
  repeat(n, function(i) {
    var whichClass = randomValue(Class);
    observe(classFeatures, whichClass);
  });
};
\end{lstlisting}
}

@@

  \subsection{Factor analysis}
{\small
\begin{lstlisting}

@@

return function() {
  return repeat(n, function(i) {
    var factors = randomValue(Factors);
    observe(getPoint, factors);
  });
};
\end{lstlisting}
}

@@

  \subsection{Hidden Markov model}
{\small
\begin{lstlisting}

@@

return function() {
  return repeat(n, function(sampIndex) {
    observeStates(randomValue(State), 0);
  });
};
\end{lstlisting}
}

@@

\subsection{Latent Dirichlet allocation}
{\small
\begin{lstlisting}

@@

\end{lstlisting}
}

@@

{\small
\begin{lstlisting}

@@

return function() {
  return repeat(inputs.length, function(sampIndex) {
    var hiddenLayer = inputToHidden(inputs[sampIndex]);
    observe(hiddenToOutput, hiddenLayer);
  });
};
\end{lstlisting}
}
